Fixing Vulnerabilities Hinders Maintainability

### 2019-02-01 paper/ICPC19/icpc19.tex
conclusion up

>>>>

This work presents an empirical study on the impact of 607 security refactorings on
the maintainability of 94 open-source projects. We leveraged Better Code Hub reports
of each project version to calculate maintainability based on a model proposed in 
previous work.

Security refactorings significantly affect the software maintainability of open-source 
software, i.e., developers hinder maintainability when fixing vulnerabilities. We found
this evidence in 40.2\% of the studied cases. Regarding regular commits, it seems that
overall maintainability increases (50.16\%). However, there is still a very considerable amout of
cases that harm maintaniability (41.35\%). This study also exhibits evidence that some patterns
need more attention than others. Broken Autentication issues, Denial-of-Service attacks 
and Memory Leaks endure maintainability in 55.3\%, 51.3\% and 47.5\%, respectively.

Following the implications of our work, we draw attention for the need of secure security frameworks,
code best practices documentation to produce safe and maintainable code, proper authorization 
mechanisms design and implementation, better tools to introduce in the CI/CD pipeline to help software
maintenance and new or improved programming languages to allow developers to easily implement safety in
their software without endangering maintainability.

As future work, this study may be extended in many directions such as analyze which guidelines are
more likely to affect security refactorings; evaluate the impact of programming languages of the
maintainability of each refactoring pattern; investigate if there is any correlation between 
software maintainability and projects popularity; expand our methodology with other software quality
properties; and, validate these findings with closed or private software.
